A New Metric for the Analysis of Swarms using Potential Fields

### 2018-08-07 paper.tex
Updated
Minor

%\section{Introduction\label{methods:SwarmStability}}
\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm, this movement is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. Measuring this unnecessary movement allows the effectiveness of a swarming algorithm to be evaluated. If the internal movement can be reduced then the stability of the swarm is improved and more of the swarms energy is focused on the goal. Reducing the internal movement also provides a more stable platform for the deployment of sensor arrays. Both the metrics presented in this paper allow this evaluation to be carried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent vary producing irregular inter-agent distances even though the interactions at the vector level indicate that the agents are in fact stable and optimum. This paper discusses the two metrics and how they differ in their approach to identify these changes. \textit{Distance-based} metrics uses variations in the inter-agent spaces as discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}. The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are defined by interactions of an agent with its neighbours this allows the ``degree'' of a relationship to be evaluated globally and therefore allows a swarm to be identified as expanding or contracting.

@@

The remainder of this paper is structured as follows Section~\ref{Section:Related} discusses different metrics that are currently used to analyse swarms, Section~\ref{Section:SwarmModelling} introduces a basic swarm model to produce an environment that can be analysed and evaluated using the two metrics. Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect and how the potential fields are applied. Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement. Section~\ref{Section:StabilityNullVector} discusses the effect of the model on swarm stability and the effects models have on swarm movements. Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis. Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric and how it is calculated and applied. Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area. Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discusses additional work that needs to be carried out to exploit the additional information that the new metric provides.

>>>>

%\section{Introduction\label{methods:SwarmStability}}
\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved such as those produced by Reynolds~\cite{REY:87}. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm, this movement is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. Measuring this unnecessary movement allows the effectiveness of a swarming algorithm to be evaluated. If the internal movement can be reduced then the stability of the swarm is improved and more of the swarms energy is focused on the goal. Reducing the internal movement also provides a more stable platform for the deployment of sensor arrays. Both the metrics presented in this paper allow this evaluation to be carried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent vary producing irregular inter-agent distances even though the interactions at the vector level indicate that the agents are in fact stable and optimum. This paper discusses the two metrics and how they differ in their approach to identify these changes. \textit{Distance-based} metrics uses variations in the inter-agent spaces as discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}. The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are defined by interactions of an agent with its neighbours this allows the ``degree'' of a relationship to be evaluated globally and therefore allows a swarm to be identified as expanding or contracting.

@@

The remainder of this paper is structured as follows Section~\ref{Section:Related} discusses different metrics that are currently used to analyse swarms, Section~\ref{Section:SwarmModelling} introduces a basic swarm model to produce an environment that can be analysed and evaluated using the two metrics. 
Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect and how the potential fields are applied. 
Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement. Section~\ref{Section:StabilityNullVector} discusses the effect of the model on swarm stability and the effects models have on swarm movements. 
% Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. 
Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis. Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric and how it is calculated and applied. Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area. Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discusses additional work that needs to be carried out to exploit the additional information that the new metric provides.
